---
title: 'Workflow 3: Two Scale Analysis of Diversity Components'
output: html_document
---

objective:
to compare diversity components at local and group scales

input:
mobr object created from original data (step 1)

output:
Figures showing species richness and rarefied richness for different types/treatments across scale. Number of species on y axis and type/treatment on x, compared for local and group level. 
Figure showing number of individuals across treatments, at local and group scale. 



```{r}
head(mobr_object)

twoscale_output = get_mob_stats(mobr_object, group = "type/treatment name", ref_group = "reference type/treatment", n_perm = 1000)
# note- some packages prevent movr package from working. Possibly dplyr. Will give message "Error in UseMethod("anova") : no applicable method for 'anova' applied to an object of class “list"". 


plot(twoscale_output, multi_panel = TRUE)

```



Now let's examine what's actually happening in this code:
```{r}

# create function. the first section is defining input and default settings. 

#It requires you to give input  of a mobr object 'mob_in'; a variable that distinguishes levels or treatments between your samples 'group_var'; and a reference group to compare against, so that's one level in your treatment 'ref_group'. 
# The function specifies those metrics it will look at, and the defaults for number of rarefied samples, miimum, number of permutations and confidence level to to calculate significance.
function (mob_in, group_var, ref_group = NULL, index = c("N", 
    "S", "S_rare", "S_asymp", "S_PIE"), n_rare_samples = NULL, 
    n_rare_min = 5, n_perm = 200, boot_groups = F, conf_level = 0.95) 

#  
  {
 # this code gives an error message if you do not provide sufficient n of permutations to run. it completely stops the function.
     if (n_perm < 1) 
        stop("Set nperm to a value greater than 1")
  # create an object called indices containing the names  below 
   INDICES <- c("N", "S", "S_rare", "S_asymp", "PIE", "S_PIE")
  #
    index <- match.arg(index, INDICES, several.ok = TRUE)
  # creates object group_id which contains the mobr_object environmental variable for the column that you specified as grouping variable. Turns this into a factor
      group_id = factor(mob_in$env[, group_var])
  #
        if (is.null(ref_group)) 
        ref_group = levels(group_id)[1]
    if (!ref_group %in% levels(group_id)) 
        stop("ref_group has to be one level in group_var!")
    group_id = relevel(group_id, ref_group)
    group_bin = factor(rep("control", times = length(group_id)), 
        levels = c("control", "treatment"))
    group_bin[group_id != ref_group] <- "treatment"
    print(index)
    samples_N = rowSums(mob_in$comm)
    samples_per_group <- table(group_id)
    if (any(is.null(n_rare_samples)) | !is.numeric(n_rare_samples)) {
        N_min_sample = min(samples_N)
        n_rare_samples = N_min_sample
    }
    else {
        n_rare_samples <- floor(n_rare_samples)
    }
    if (any(n_rare_samples < n_rare_min)) {
        warning(paste("The number of individuals for rarefaction analysis is too low and is set to the minimum of", 
            n_rare_min, "individuals."))
        n_rare_samples[n_rare_samples < n_rare_min] <- n_rare_min
        n_rare_samples <- unique(n_rare_samples)
        print(paste("Number of individuals for rarefaction:", 
            paste(n_rare_samples, collapse = ", ")))
    }
    n_rare_groups <- n_rare_samples * min(samples_per_group)
    abund_group = aggregate(mob_in$comm, by = list(group_id), 
        FUN = "sum")
    dat_groups <- calc_biodiv(abund_mat = abund_group[, -1], 
        groups = abund_group[, 1], index = index, n_rare = n_rare_groups)
    dat_samples <- calc_biodiv(abund_mat = mob_in$comm, groups = group_id, 
        index = index, n_rare = n_rare_samples)
    if (any(index == "S")) {
        gamma <- with(dat_groups, value[index == "S"])
        alpha <- with(dat_samples, value[index == "S"])
        beta_S <- gamma[group_id]/alpha
        beta_S[!is.finite(beta_S)] <- NA
        dat_betaS <- data.frame(group = group_id, index = "beta_S", 
            n_rare = NA, value = beta_S)
        dat_samples <- rbind(dat_samples, dat_betaS)
    }
    if ("S_rare" %in% index) {
        for (i in 1:length(n_rare_samples)) {
            gamma <- with(dat_groups, value[index == "S_rare" & 
                n_rare == n_rare_groups[i]])
            alpha <- with(dat_samples, value[index == "S_rare" & 
                n_rare == n_rare_samples[i]])
            beta_S_rare <- gamma[group_id]/alpha
            beta_S_rare[!is.finite(beta_S_rare)] <- NA
            dat_beta_S_rare <- data.frame(group = group_id, index = "beta_S_rare", 
                n_rare = n_rare_samples[i], value = beta_S_rare)
            dat_samples <- rbind(dat_samples, dat_beta_S_rare)
        }
    }
    if ("S_asymp" %in% index) {
        gamma <- with(dat_groups, value[index == "S_asymp"])
        alpha <- with(dat_samples, value[index == "S_asymp"])
        beta_S_asymp <- gamma[group_id]/alpha
        beta_S_asymp[!is.finite(beta_S_asymp)] <- NA
        dat_beta_S_asymp <- data.frame(group = group_id, index = "beta_S_asymp", 
            n_rare = NA, value = beta_S_asymp)
        dat_samples <- rbind(dat_samples, dat_beta_S_asymp)
    }
    if ("S_PIE" %in% index) {
        gamma <- with(dat_groups, value[index == "S_PIE"])
        alpha <- with(dat_samples, value[index == "S_PIE"])
        beta_S_PIE <- gamma[group_id]/alpha
        beta_S_PIE[!is.finite(beta_S_PIE)] <- NA
        dat_beta_S_PIE <- data.frame(group = group_id, index = "beta_S_PIE", 
            n_rare = NA, value = beta_S_PIE)
        dat_samples <- rbind(dat_samples, dat_beta_S_PIE)
    }
    F_obs <- get_F_values(dat_samples, permute = F)
    F_rand <- dplyr::bind_rows(replicate(n_perm, get_F_values(dat_samples, 
        permute = T), simplify = F)) %>% ungroup()
    F_obs <- F_obs %>% mutate(F_val_obs = F_val, F_val = NULL)
    F_rand <- left_join(F_rand, F_obs)
    p_val_samples <- F_rand %>% group_by(index, n_rare) %>% summarise(p_val = sum(F_val_obs <= 
        F_val)/n_perm) %>% ungroup()
    if (!boot_groups) {
        diff_obs <- get_group_diff(mob_in$comm, group_bin, index, 
            n_rare = n_rare_groups, permute = F)
        diff_rand <- bind_rows(replicate(n_perm, get_group_diff(mob_in$comm, 
            group_bin, index, n_rare = n_rare_groups, permute = T), 
            simplify = F))
        diff_obs <- diff_obs %>% mutate(d_obs = delta, delta = NULL)
        diff_rand <- left_join(diff_rand, diff_obs)
        p_val_groups <- diff_rand %>% group_by(index, n_rare) %>% 
            summarise(p_val = get_pval(rand = delta, obs = first(d_obs), 
                n_samples = n_perm)) %>% ungroup()
    }
    else {
        abund_dat <- cbind(group_id, mob_in$comm)
        boot_repl_groups <- replicate(n_perm, boot_sample_groups(abund_dat, 
            index = index, n_rare = n_rare_groups), simplify = F)
        boot_repl_groups <- bind_rows(boot_repl_groups)
        alpha <- 1 - conf_level
        p <- c(alpha/2, 0.5, 1 - alpha/2)
        dat_groups <- boot_repl_groups %>% group_by(group, index, 
            n_rare) %>% do(setNames(data.frame(t(quantile(.$value, 
            p, na.rm = T))), c("lower", "median", "upper")))
    }
    dat_samples$index <- factor(dat_samples$index, levels = c("N", 
        "S", "beta_S", "S_rare", "beta_S_rare", "S_asymp", "beta_S_asymp", 
        "PIE", "S_PIE", "beta_S_PIE"))
    dat_samples <- dat_samples[order(dat_samples$index, dat_samples$n_rare, 
        dat_samples$group), ]
    dat_groups$index <- factor(dat_groups$index, levels = index)
    dat_groups <- dat_groups[order(dat_groups$index, dat_groups$n_rare, 
        dat_groups$group), ]
    dat_samples$index <- factor(dat_samples$index)
    if (!boot_groups) {
        p_val_groups$index <- factor(p_val_groups$index, levels = index)
        p_val_groups <- p_val_groups[order(p_val_groups$index), 
            ]
        out <- list(samples_stats = dat_samples, groups_stats = dat_groups, 
            samples_pval = p_val_samples, groups_pval = p_val_groups, 
            p_min = 1/n_perm)
    }
    else {
        out <- list(samples_stats = dat_samples, groups_stats = dat_groups, 
            samples_pval = p_val_samples, p_min = 1/n_perm)
    }
    class(out) = "mob_stats"
    return(out)
}
<bytecode: 0x11f3b27f8>
<environment: namespace:mobr>



```





